PM Mungbean meta-analysis

### 2020-09-01 02_Preliminary_analysis.Rmd
Label changes

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra)
source("R/import_data.R")
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
```

## Explore and Visualise Data
First let's import the data.
Each row provides the mean yield for each experimental treatment.

```{r echo=TRUE, message=FALSE, warning=FALSE}
PM_MB_means <- import_data() 
```

The data explored in the `PM_MB_means` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will focus on the change in mungbean yields, which is attributable to the fungicide treatment. 
Later in this document we will also consider the impact of fungicide treatments on disease incidence/severity.
Mean powdery mildew incidence was recorded at the end of each growing season.
Some experiments recorded disease incidence throughout the season and these were summarised with AUDPC.
AUDPC for trials which only provided final powdery mildew incidence was calculated using the record of 'first sign of disease' and final assessment.
More details on how the `PM_MB_means` data was compiled can be found in the ['Data wrangle vingette'](./vignette/191115_PMMB_DataWrangling_PM.html)

@@

   8 In-crop rainfall

   9 Irrigated versus non-irrigated trials

   10 Mean daily temperature in the weeks following spray application

   11 Final disease rating

   12 Two pathogen species cause PM (new discovery in 2018)

The data plotted below are from `r length(unique(PM_MB_means$trial_ref))` field trials between (`r first(levels(PM_MB_means$year))` - `r last(levels(PM_MB_means$year))`) of which the details are described in the [studies considered in meta-analysis section](#studies-considered-in-meta-analysis).

@@

First let's standardise the sulphur treatments.
Some early experiments tested different formulations.
There was no difference in the sulphur formulations so we will refer to all of them as "sulphur".

```{r sulphur_simplify}
# Remove the variation in different sulphur formulations
unique(PM_MB_means[grep("sulphur", PM_MB_means$fungicide_ai),
                   "fungicide_ai"])
PM_MB_means[grep("sulphur", PM_MB_means$fungicide_ai), "fungicide_ai"] <-
   "sulphur"
```

Let's explore how many different types of fungicide were used and at what frequency.

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   coord_flip()
```

The demethylation inhibitors (DMI), tebuconazole and propiconazole, are used in the highest frequencies, along with sulphur.
The DMIs have the same fungicide mode of action and are good candidates to be pooled in the meta-analysis.

Amistar Xtra and Custodia both contain strobilurin and triazole, however, because they contain differing dose ratios (inverted) pooling may not be appropriate.

Perhaps best way forward is to focus the meta-analysis on only the DMIs.
This can then be compared to an additional meta-analysis including azoxystrobin as a comparison.

@@

Let's investigate the difference in the effect of the type of fungicide on proportion of yield saved, but first we need to calculate yield saved relative to the no spray control.

```{r Calculate_yield-savings}
PM_MB_means$grain_yield.t.ha. <-
   as.numeric(PM_MB_means$grain_yield.t.ha.) # NAs produced due to some cells having text description to why there is no specific data

for (i in unique(PM_MB_means$trial_ref)) {
   # Loop code over each trial reference
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]

   if (any(is.na(unique(dat1$row_spacing)))) {
      # If there is no recorded row.spacing don't calculate proportional yield, skip to the next trial
      warning(
         unique(dat1$trial_ref),
         "at",
         unique(dat1$location),
         "in year",
         unique(dat1$year),
         "has unknown row.spacing (NA)\n 'yield_gain' and 'proportional yield gain' not calculated\n"
      )
      next()
   }

      # Reassign values to row_spacing subset
      dat1[dat1$row_spacing == j, ] <- dat2
   }
   # Reassign values to trial_ref subset
   PM_MB_means[PM_MB_means$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data
dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)
PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

write.csv(PM_MB_means,
          "cache/1911_PM_MB_means&Ygains.csv")
```

Now we have calculated proportional yield savings we can inspect the difference between the fungicide active ingredients

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_means %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

**  **  *

```{r Fungicide_x_nSprays_tble1, message=FALSE, warning=FALSE}
PM_MB_means %>%
   mutate(fungicide_ai = as.factor(PM_MB_means$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole + 120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by fungicide type") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.
Interestingly sulphur sprays show good protection when sprayed twice.

#### Fungicide Doses  
From this point on we will investigate only the DMI fungicides, tebuconazole and propiconazole, and disregard all other fungicide modes-of-action.

@@

```{r spray_relative2FS}
PM_MB_means %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>%
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(labels = c(
      "First spray",
      "Second spray",
      "Third spray",
      "Fourth spray",
      "Fifth spray"
   )) +
   ylab("Days to First Sign of Disease") +
   facet_grid(cols = vars(total_fungicide)) +
   coord_flip()
```

The plot shows single fungicide applications are mostly made at zero, first sign of disease in the crop.
Experiments that tested two spray treatments also tend to make the first application at first sign of disease, and the second spray between 10 and 25 days later.
Three spray treatments seemed to be mostly used in experiments where preventative sprays as the first fungicide application were made.

@@

```{r Host_genotype}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

#### Genotype yield variability

Does host cultivar significantly impact the yield?
Let's look at the volatility in yields to see if any cultivar yields better or has greater tolerance to powdery mildew.
Below we plot distributions of mean trial yields for each cultivar.
The following plots use data that includes all fungicide types.

First let's look at overall yields for each cultivar, where trial is not a factor, for all treatments (spray and no spray).

```{r yield_vol}
# Outputs the table and graph
source("R/yield_volatility.r") # function to investigate the volatility in yields
# Volatility is the range between the upper and lower 97.5 and 2.5 % quartiles
YV1 <-
   yield_volatility(genotype_by_trial = FALSE, control_only = FALSE)
YV1[[2]] + 
   scale_fill_usq()
YV1[[3]]
```

This plot shows a large amount of yield volatility in all cultivars except Green Diamond.
Green Diamond shows little volatility probably because this cultivar was only used in `r length(unique(PM_MB_means[PM_MB_means$host_genotype == "Green Diamond","trial_ref"]))` trial.

Now let's look at overall yields for each cultivar, where trial is not a factor, in only the no spray control treatments.

```{r yieldVol_control}
YV2 <-
   yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
YV2[[2]] +
   scale_fill_usq()
YV2[[3]]
```

Volatility of the cultivars are less, albeit fairly similar, in control plots when trial is not considered as a factor.

Let's examine the same plots as above but only for the most commonly used trial site, Hermitage.

```{r yieldVol_Herm}
YV1_herm <-
   yield_volatility(
      genotype_by_trial = FALSE,
      control_only = FALSE,
      location = "Hermitage"
   )
YV1_herm[[2]] +
   scale_fill_usq()

YV1_herm[[3]]
```

Jade shows a large amount of volatility, excluding Green Diamond, Crystal shows the next least volatility.

Let's take a closer look at Crystal with spray management treatments and without.

```{r YieldVol_Crystal}
YV1_crystal <-
   yield_volatility(genotype = "Crystal", control_only = FALSE)
YV1_crystal[[2]] +
   scale_fill_usq()

YV1_crystal[[3]]

YV2_crystal <-
   yield_volatility(genotype = "Crystal", control_only = TRUE)
YV2_crystal[[2]] +
   scale_fill_usq()

YV2_crystal[[3]]
```

There is a lot of variation between field trials for Crystal, but little variation within field trials. This indicates that trial should be fit as a random effect when we fit a model to this data.

Let's look at Jade, to see if it differs.

```{r YieldVol_Jade}
YV1_jade <- yield_volatility("Jade",control_only = FALSE)
YV1_jade[[2]] +
   scale_fill_usq()

YV1_jade[[3]]
```

For Jade there is high variation between trials and with-in.
Perhaps indicating lower tolerance. 
Let's look at the same plot with data only from no spray controls to see if the variation is the same.

```{r YieldVol_JadeControl}
YV2_jade <- yield_volatility("Jade", control_only = TRUE)
YV2_jade[[2]] +
   scale_fill_usq()

YV2_jade[[3]]
```

There still remains a higher volatility in Jade and perhaps a greater shift in the probability curve including fungicide treatments versus probability curves including only no spray controls.
Indicating that there might be a difference in powdery mildew tolerance between Jade and Crystal.

Overall there seems to be an influence of cultivar.
However this variation is not as distinct as the variation between trials.

@@

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.1) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

### Yield vs in-season rain

In0-crop rain could be a significant factor which contributes to the end of season grain yield.
To investigate this effect I will use a linear model (GY ~ SR) where seasonal rainfall represents a specific time frame with respect to planting time.
Multiple time frames were tested from 20 days prior to planting, to 100 days after planting.

The multiple time frames were tested by a loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall.
The code for this process was moved to be contained within it's own script and can be run as a job as it can take a long time.
To speed up the computation I split the script into four, ran the jobs and bound the data into one frame.
The output of the script is saved with the prefix `lmInSeasonRainfall` followed by the time windows the model is run.
For example `lmInSeasonRainfall_20.40_50.80.csv` holds the model results for start days between 20 and 40 days after planting to end window dates between 50 and 80 days after planting.

This tests when rainfall during the cropping season could be best associated with yield

```{r rainfall_sum_Sstart}
# The four data files were created from four scripts run as four jobs in RStudio
# when the four jobs are finished they can be imported and combined into one data.frame
# source(Rainfall_x_cropYield_s-10.09_e30.65.R)
# source(Rainfall_x_cropYield_s-10.09_e65.100.R)
# source(Rainfall_x_cropYield_s10.30_e30.65.R)
# source(Rainfall_x_cropYield_s10.30_e65.10.R)

lm_rain1 <-
   read.csv("cache/lmInSeasonRainfall_-10.09_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain2 <-
   read.csv("cache/lmInSeasonRainfall_-10.09_65.100.csv",
            stringsAsFactors = FALSE)
lm_rain3 <-
   read.csv("cache/lmInSeasonRainfall_10.30_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain4 <-
   read.csv("cache/lmInSeasonRainfall_10.30_65.100.csv",
            stringsAsFactors = FALSE)

lm_rain <- rbind(lm_rain1, lm_rain2, lm_rain3, lm_rain4)
```

Now we have read in the data from the previously computed models let's find the linear model which has:  
   - The lowest P value  
   - the highest adjusted r squared value

```{r WhichLModel}
lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ]
lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ]
lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)), ]
```

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows.
log of the p value was used to improve the resolution of small very small p values.

```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval))) +
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu", direction = -1) 
```

```{r rainLMadjrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_adj_rsquared)) +
   geom_raster(aes(fill = lm_adj_rsquared), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu") 
```

The approximate period when in-crop rainfall affects the most on the mean harvest is between 14 - 67 days after planting.

### Season range and first incidence

Mungbean is a summer crop and powdery mildew requires cool conditions to establish.
Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons.
Can we visualise when PM is most likely to establish over the calendar year?

I'll build a plot of horizontal lines indicating the season length and use a dot on the line to indicating the day of the year first sign was observed for that trial.

```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(
      Planting_date = unique(planting_date),
      Harvest_date = unique(harvest_date),
      First_sign_PM = unique(first_sign_disease),
      season_length = as.Date(unique(harvest_date)) - as.Date(unique(planting_date)),
      cropping_season = "Actual"
   )

# For some trials where planting date and Harvest date is missing let's estimate 
#     the season length by using the mean of all season lengths
season_dates[is.na(season_dates$season_length),"cropping_season"] <- "Estimated"

season_dates[is.na(season_dates$Planting_date),"Planting_date"] <- 
   season_dates[is.na(season_dates$Planting_date),"Harvest_date"] - 
   as.numeric(mean(season_dates$season_length, na.rm = TRUE))

season_dates[is.na(season_dates$Harvest_date),"Harvest_date"] <- 
   season_dates[is.na(season_dates$Harvest_date),"Planting_date"] + 
   as.numeric(mean(season_dates$season_length, na.rm = TRUE))

season_dates[is.na(season_dates$season_length),"season_length"] <- 
   mean(season_dates$season_length, na.rm = TRUE)

season_dates$Planting_day <- yday(season_dates$Planting_date)
season_dates$Disease_day <- yday(season_dates$First_sign_PM)

season_dates[c(1:2,6), "Planting_day"] <- 1

season_dates$Harvest_day <-
   season_dates$Planting_day + season_dates$season_length

first_day_month <- c(0, 31, 60, 91, 121, 152, 182, 212, 244, 274) + 1
axis_labels_date <-
   c(
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[1], "%b-%d"),
      # Jan
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[2], "%b-%d"),
      # Feb
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[3], "%b-%d"),
      # March
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[4], "%b-%d"),
      # April
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[5], "%b-%d"),
      # May
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[6], "%b-%d"),
      # June
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[7], "%b-%d"),
      # July
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[8], "%b-%d"),
      # August
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[9], "%b-%d"),
      # Sept
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[10], "%b-%d") # October
   )

>>>>

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               here)

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
load("cache/ImportDataAndSelectTrials01.Rdata")
```

## Explore and Visualise Data

The data explored in the `PM_MB_dat` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will describe the data and then investigate any change in mungbean yields which is attributable to the fungicide treatment.

### Descriptive statistics  
#### Spatio-temporal variation

Lets look at how many trials were undertaken each season

```{r FinalTrialsForInclusion1}
tmp1 <- PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   distinct(trial_ref, location, year)%>%
   arrange(year)
   table(tmp1$year)
```

Also the location of the trials

```{r FinalTrialsForInclusion}
   table(tmp1$location)
```

#### Mean yield

What is the variation in yield over all the trials (including non-DMI fungicide trials) in the no spray control and fungicide treated plots.

```{r ControlYield_AllTrials}
PM_MB_dat %>%
   filter(fungicide_ai == "control")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.),
             minimum_yield = min(grain_yield.t.ha.),
             maximum_yield = max(grain_yield.t.ha.),
             yield_stdev = sd(grain_yield.t.ha.)) %>%
   kable(align = 'c',
   caption = "Average yield statistics for no spray contol plots"
)

What is the yield variation in no spray control trials yield over only the DMI fungicide trials.  
```{r ControlYield_DMITrials}
DMITrials <- PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   distinct(trial_ref)%>%
   pull()

PM_MB_dat %>%
   filter(trial_ref %in% DMITrials)%>%
   filter(fungicide_ai == "control" )%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))%>%
   kable(align = 'c',
   caption = "Average yield statistics for control plots in trials which used a DMI fungicide")

Mean yields in DMI fungicide treated plots
```{r Yield_DMITrials}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   summarise(mean_yield = mean(grain_yield.t.ha.*1000),
             minimum_yield = min(grain_yield.t.ha.*1000),
             maximum_yield = max(grain_yield.t.ha.*1000),
             yield_sterr = sd(grain_yield.t.ha.*1000)/(sqrt(length(grain_yield.t.ha.))))%>%
   kable(align = 'c',
   caption = "Average yield statistics for DMI fungicide treated plots")

Lets look at the yield gains in each trial compared to the no spray control, we will also calculate the proportional grain yield gains.

```{r Calculate_yield-savings}
for (i in unique(PM_MB_dat$trial_ref)) {
   # Loop code over each trial reference
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]

   if (any(is.na(unique(dat1$row_spacing)))) {
      # If there is no recorded row.spacing don't calculate proportional yield, skip to the next trial
      warning(
         unique(dat1$trial_ref),
         " at ",
         unique(dat1$location),
         " in year ",
         unique(dat1$year),
         " has unknown row.spacing (NA)\n 'yield_gain' and 'proportional yield gain' not calculated\n"
      )
      next()
   }

      # Reassign values to row_spacing subset
      dat1[dat1$row_spacing == j, ] <- dat2
   }
   # Reassign values to trial_ref subset
   PM_MB_dat[PM_MB_dat$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data
dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ]

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)

PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ] <- dat1
```

Lets show the yield gains in each trial as a table and plot
```{r YieldGain_table}
PM_MB_dat %>%
   filter(trial_ref %in% DMITrials)%>%  # selects only the trials with DMI fungicide treatments
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   group_by(trial_ref, location, year)%>%
   summarise(yield_gain_m = mean(yield_gain),
             yield_gain_se = sd(yield_gain, na.rm = T)/sqrt(n()),
             yg_n = n(),
             .groups = "drop")%>%
   arrange(yield_gain_m)
```

```{r YieldGain_plot}
PM_MB_dat %>%
   filter(trial_ref %in% DMITrials)%>%  
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   group_by(trial_ref, location, year)%>%
   summarise(yield_gain_m = mean(yield_gain),
             yield_gain_se = sd(yield_gain, na.rm = T)/sqrt(n()),
             .groups = "drop")%>%
   mutate(trial = paste(location,year,sep = "\n"))%>%
   ggplot(aes(reorder(trial, yield_gain_m), yield_gain_m))+
   geom_bar(stat = "identity")+
   geom_errorbar(width = 0.3 ,aes(ymax = yield_gain_m + yield_gain_se, 
                     ymin = yield_gain_m - yield_gain_se))+
   theme(axis.text.x = element_text(angle = 90, hjust = 1))

@@

   8 Final disease rating / disease pressure

The data plotted below are from `r length(unique(PM_MB_dat$trial_ref))` field trials between (`r first(levels(PM_MB_dat$year))` - `r last(levels(PM_MB_dat$year))`) of which the details are described in the [Trials considered for inclusion in meta-analysis section](#Trials-considered-for-inclusion-in-meta-analysis).

<br>
<br>
<br>

@@

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_dat %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

<br>
<br>
<br>

```{r Fungicide_x_nSprays_tble1, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   mutate(fungicide_ai = as.factor(PM_MB_dat$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole + 120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\nfaceted by number of sprays") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.

@@

#### Fungicide Doses{#fungicide-doses}  
From this point on we will investigate only the DMI fungicides, tebuconazole and propiconazole, and disregard all other fungicide modes-of-action.

@@

```{r spray_relative2FS}
PM_MB_dat %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>%
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(labels = c(
      "First spray",
      "Second spray",
      "Third spray",
      "Fourth spray",
      "Fifth spray"
   )) +
   ylab("Number of days after or before first sign of powdery mildew") +
   ggtitle("Fungicide application timing relative to First Sign of Disease.\n - Faceted by number of applications")+
   facet_grid(cols = vars(total_fungicide)) +
   coord_flip()
```

The plot shows single fungicide applications are mostly made at zero, first sign of disease in the crop. Some were made earlier and some around 10 days later.  
Experiments that tested two spray treatments also tend to make the first application at first sign of disease, and the second spray between 10 and 25 days later.
Three spray treatments seemed to be mostly used in experiments where preventative sprays as the first fungicide application were made.

@@

```{r Host_genotype}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

@@

Lets plot disease severity in boxplots grouped by row spacing treatment.
We are asking here, which row spacing leads to the higher disease severity?

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew severity between different row spacing across all trials")
```

There is little difference here between the row spacing treatments, lets look if there is a trial location effect as disease pressure lead to varying results.

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.05) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

@@

<br>
<br>
<br>
<br>

### Yield loss in cultivar Jade

Mungbean Jade shows the best quantitative resistance to powdery mildew.
Based on a GRDC report by Sue Thompson [-@SueThompson2016] we want to know what the yield loss was in on cultivar Jade in the 2016 Trials.
```{r JadeYieldLoss}
PM_MB_dat %>%
   filter(year == "2016",
          host_genotype == "Jade")%>%
   mutate(Treatment = case_when(
      fungicide_ai == "control" ~ "control",
      TRUE ~ "FungicideTreated"))%>%
   group_by(location, Treatment)%>%
   summarise(trial_ref = trial_ref,
             location = location, 
             year = year,
             Yield = mean(grain_yield.t.ha., na.rm = TRUE),
             .groups = "drop")%>%
   distinct()%>%
   pivot_wider(names_from = Treatment, values_from = Yield)%>%
   mutate(Diff = FungicideTreated - control) %>%
   mutate(percent = (Diff/FungicideTreated)*100)%>%
   select(trial_ref, location, year, control, FungicideTreated,percent, Diff)
